---
title: "Prediction of manner in which personnal activity is done very well"
author: "Nabié Binouni"
output: html_document
---

# Abstract

The aim of this analysis is to predict personnal activity well done using machine learning. The variable of interest is **classe**. The random forest algorithm is used to train the model and missclassify one sample on the validation dataset. I've predicted 20 sample for sumitting in Coursera.
  
# Analysis

The dataset used for this analysis is from <http://groupware.les.inf.puc-rio.br/har>.
It's about: Six young health participants were asked to perform one set of 10 
repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

### Dataset loading
Two datasets are used for this analysis: **pml_train** for performing the model 
and **pml_test** for prediction.The variable of interest is **classe** which is factor variable. So it is a classification issue. As it has five levels, we can't use logistic regression. The first model that comes in mind is linear discriminant analysis. But I'm going to use a stacking model because of accuracy. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```
```{r load, message=FALSE}
library(caret)

library(randomForest) # rf

train_url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

if(!file.exists("pml-training.csv")){ 
    download.file(train_url, destfile = "pml-training.csv")
    }

if(!file.exists("pml-testing.csv")){ 
    download.file(test_url, destfile = "pml-testing.csv")
}

pml_train <- read.csv("pml-training.csv", na.strings = c("", "NA"))
pml_test <-  read.csv("pml-testing.csv", na.strings = c("", "NA"))

cbind(freq = table(pml_train$classe), percentage = round( prop.table(table(pml_train$classe))*100, digits = 2))

```
### Dataset cleaning

The variable with a lot of missing values are going to be removed.
There are 100.
```{r clean}
#SumNA <- function(x) sum(is.na(x))
dim(pml_train[sapply(pml_train, function(x) sum(is.na(x)) == 19216)])
training <- pml_train[sapply(pml_train, function(x) sum(is.na(x)) != 19216)]
testing <- pml_test[sapply(pml_test, function(x) sum(is.na(x)) != 19216)]

```

### Trainind and validation datasets

For the training purpose, the dataset is splitted into **train** 70% and **validation** datasets. This last one will allow to detemine the model accuracy.
```{r Split}
set.seed(245)
inTrain <-  createDataPartition(training$classe, p = .7, list = FALSE)
train <-  training[inTrain, ]
validation <-  training[-inTrain, ]

```

### Model
I've tried some algorithm not shown here such as lda, svm, knn and cart. The accuracy of stacking this models is less than the rf one's. So the random forest algorithm is used to train the model becaus of it's accuracy. But this algorithm is known for it's overfitting.To avoid this, k-fold cross-validation is used.
```{r Model, cache = TRUE}

seed <- 1235
traincontrol <-  trainControl(method = "cv", number = 10)

# rf
set.seed(seed)
rf <- train(classe ~., data = train, method = "rf",
           trControl = traincontrol, metric = "Accuracy")

#print(rf)
```

### Accuracy
Here we determine the accuracy of the model, then the out error rate.

```{r Accuracy}
rf_predict <- predict(rf, validation)
rf_accuracy <- confusionMatrix(rf_predict, validation$classe)
print(rf_accuracy)
```
The model accuracy is **`r rf_accuracy$overall[1]`**. So out error rate is **`r format(1-rf_accuracy$overall[1])`**.
It misclassified one sample. We can claim that the random forest algorithm is the best one for this issue.  

### prediction on testing dataset
Here are the results of prediction for 20 samples, to submit.
```{r prediction}
classe <- predict(rf, testing)
write.csv(as.data.frame(classe), file = "Prediction.csv", row.names = FALSE)
```

